---
title: "legtext"
description: |
  A short description of the post.
author:
  - name: David L. Passmore
    url: https://davidpassmore.net
    orcid_id: 0000-0001-9307-5164
    affiliation: Distinguished Professor, Emeritus, Penn State; Academic Visitor, University of Pittsburgh
categories:
  - Act 16
  - Act 85
  - Beaver County, Pennsylvania 
  - ethane
  - ethane cracker
  - ethylene
  - Pennsylvania legislature
  - Pennsylvania Petrochemicals Complex
  - Shell Chemical Appalachia LLC
  - Shell, plc
  - tax expenditures
date: 2022-8-10
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  markdown: 
    wrap: 72
draft: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.

## Last updated on {.appendix}

```{r,echo=FALSE}
Sys.time()
```

## Your Comments & Corrections {.appendix}

To make comments about this posting or to suggest changes or
corrections, send email to [David
Passmore](mailto:dlp@davidpassmore.net), send a direct message on
Twitter at [\@DLPPassmore](https://twitter.com/dlppassmore), or send an
IMsg or SMS to [dlp\@psu.edu](mailto:dlp@psu.edu){.email}.

## Reuse {.appendix}

Text and illustrations are licensed under Creative Commons Attribution
[CC BY 4.0](https://creativecommons.org/licenses/by-sa/4.0/). Source
code is available at <https://github.com/davidpassmore/blog>, unless
otherwise noted. The figures/photos/images/maps reused from other
sources do not fall under this license and are recognized by text in the
caption starting: "Figure from..." or "Photo from..." , "Image from..."
or "Map from...".
