---
title: "Cracker Impact"
description: |
  I request your answers to my questions -- for which I personally have no, or perhaps malformed, answers -- about the Pennsylvania Petrochemicals Complex in Monaca, Pennsylvania.
preview: xxx
draft: TRUE
author:
  - name: David L. Passmore
    url: https://davidpassmore.net
    orcid_id: 0000-0001-9307-5164
    affiliation: Distinguished Professor, Emeritus, Penn State; Academic Visitor, University of Pittsburgh
date: 2022-04-21
categories:
  - ethane cracker
  - shell public liability company
  - Beaver County, Pennsylvania
  - Potter Township, Pennsylvania
  - input-output modeling
  - economic impact
output:
  distill::distill_article:
    self_contained: false
    toc: true 
    toc-depth: 5
#bibliography: references.bib
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.

```{r,echo=FALSE}
Sys.time()
```

## Reuse {.appendix}

Text and figures are licensed under Creative Commons Attribution [CC BY 4.0](https://creativecommons.org/licenses/by-sa/4.0/). Source code is available at <https://github.com/davidpassmore/blog>, unless otherwise noted. The figures that have been reused from other sources do not fall under this license and are recognized by footnoted text: "Figure from ...".

## Comments/Corrections {.appendix}

To make comments about this posting or to suggest changes or corrections, send email to [David Passmore](mailto:dlp@davidpassmore.net), send a direct message on Twitter at [\@DLPPassmore](https://twitter.com/dlppassmore), or send an IMsg or SMS to [dlp\@psu.edu](mailto:dlp@psu.edu){.email}.
