---
title: "understanding input-output modeling for assessing potential economic impact"
description: |
  Maybe this eesay won't shock your soul, but describes input-output models, assumptions, & uses to estimate economic impact.
preview: types-of-economic-impact.jpg
author:
  - name: David L. Passmore
    url: https://davidpassmore.net
    orcid_id: 0000-0001-9307-5164
    affiliation: Distinguished Professor, Emeritus, Penn State; Academic Visitor, University of Pittsburgh
date: 2022-04-12
categories:
  - IMPLAN
  - Wassily Leontief
  - economic impact
  - counterfactuals
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Input-output modeling is a data-based, computational, and often
prospective approach to unwinding the impact of projects, events, and
changes in local economies. An input-output model accounts for goods and
services that are required as *inputs* by industries to produce
*outputs* of goods and services necessary to satisfy supply requirements
for inputs to other industries and delivery directly to consumers
[@miernyk1965].

Input-output modeling is pervasive in the measurement of economic
impact. However, this modeling technique often is misapplied and
debauched by various rakehells of academia and consulting to fulfill
patently politically-motivated spending of public funds. For these
reasons, understanding the ways and means of input-output modeling is
essential to weigh assertions about economic impact. Frankly, though, I
wrote this posting for more personal reasons: I abide by the apothegm
that a person knows nothing about a topic until that person had to write
about it. I will refer to this posting as a base for other postings in
which I review economic impact reports.

In this paper, I specify the details of the input-output model and its
assumptions, outline common uses (and some concerns about abuses) of the
model to estimate potential economic impact, and list selected software
and systems available for input-output modeling. In doing so, I do not
get into any depth in the vast literature on input-output modeling. Nor
do I establish a critical or skeptical stance about this modeling or its
uses. For the most part, I try to provide a detailed, yet unadorned,
description of input-output modeling.

### Specification of the Input-Output Model

#### Production and Consumption

Consider an economy with $I$ producing industries and $J$ purchasing
industries, where $I = J$. Some of the output of industry $i$ is sold to
industry $j$ for use in its own production.

Let $\textbf{X}$ stand for a square matrix of transactions among
industries, with elements $x_{ij}$ containing the dollar value of goods
and services sold by producing industry $i$ to purchasing industry $j$.
Sales of the product of industry $i$ to itself --- that is, $x_{i=j}$
--- are possible. For instance, the coal industry also might purchase
coal for use in its own production processes.

Output of industry $i$ not purchased by industries for their own
production instead is sold directly to consumers to meet demands for
personal consumption in households, government purchases of defense and
non-defense goods and services, fixed investments such as housing and
other structures, and exports net of imports. Let $\textbf{y}$ represent
an $I$--length vector whose elements display the dollar value of goods
and services delivered by producing industry $i$ to fulfill final
demands of the economy for personal consumption, government purchases,
fixed investments, and net exports. Vector $\textbf{y}$ represents the
"final demand" for goods and services in an economy and is known
commonly in the parlance of national income and product accounting
practice as *gross domestic product* [@grossdo].

#### Direct Requirements

Quantities in an input--output model most commonly are expressed as
monetary values (e.g., as dollars, rubles, yuan, or some other form of
currency used as a medium of exchange) rather than as quantities of
goods and services, although some energy and resource input--output
models are scaled in physical quantities such as megawatts of
electricity, kilograms of waste, hectares of land, or liters of water
[@weisz2006].

The production sector of the economy, then is defined by $\textbf{X}$,
and $\textbf{y}$ is the consumption sector. The total output of the
economy by producing industry, contained in $I$--length vector
$\textbf{x}$,[^1] is equal to the sum of production and consumption, or

[^1]: Although the notation seems awkward to have an element of matrix
    $\textbf{X}$ indicated by the scalar $x$ and the column vector of
    total output indicated by $\textbf{x}$, we are following notation
    conventions that are common in the field of interindustry analysis.

$$
\begin{equation}
\textbf{x}=\textbf{X}+\textbf{y}.    (1)
\end {equation}
$$

Define $\textbf{A}$, a direct requirements matrix with elements
$a_{ij}=x_{ij}/x_{j}$ showing the proportion of industry $i$ output
required for production by industry $j$. Equation (1) now becomes

$$
\begin{equation}
\textbf{x}= \textbf{Ax} + \textbf{y}. (2)
\end{equation}
$$

##### Some Assumptions About Requirements

Elements in $\textbf{A}$ are assumed to be fixed, homogeneous, and
linear within purchasing industries. I dwell on these assumptions
because they are at the very core of the operation and credibility of
input-output modeling.

*Fixed*

Inputs required by purchasing industry $j$ from producing industry $i$
do not change over time. More technically, zero elasticity of price or
quantity substitution exists among inputs to purchasing industry $j$.
Increasing or decreasing industry *j* output requires increasing or
decreasing industry *i* inputs in the same proportion. "In other words,"
according to the Bureau of Economic Research for the U.S. Department of
Commerce,[@regional] "...businesses can use as many inputs as needed
without facing higher prices." In these ways, coefficients in
$\textbf{A}$ are assumed to be *fixed*.

*Homogeneous*

*All* firms and establishments in purchasing industry $j$ use the same,
fixed proportion of industry $i$ inputs. That is, no variation exists in
resource and product markets within an industry. Same suppliers; same
customers; same technology of production. Every producer in an industry
does the same thing and is affected by market forces in the same way. In
this sense, industries are rendered *homogeneous* in input-output
modeling.

An aside about terminology and classification applicable to the the
assumption of homogeneity of industries in input-output modelings:[^2]
\n

[^2]: And to most economic analysis that uses public data to focuses on
    industry attributes.

1.  A fundamental rule followed in most input-output modeling is that
    the least common denominator of analysis occurs the level of an
    *industry*, which is an aggregation of activity by highly similar
    goods or services and business activity, and not by firm or within
    establishments run by firms. Industry output is produced by *firms*,
    which are for-profit business organizations, such as a corporations,
    limited liability companies, or partnership. A number of firms can
    contribute to an *industry*. As I shall explain in the next item in
    this side discussion, industries are not mutually exclusive
    collections of firms. Separate industries might include some of the
    same firms. Also, firms might own or manage multiple establishments
    at which industry production occurs. An *establishment* is a single
    location where business is conducted or where services/operations
    are performed.[^3] In general, the production flow moves from
    establishments within firms and, then, to industries.

2.  To add complexity to the input-output accounting scheme, some firms
    produce multiple goods and services that are classifiable in
    multiple industries. For instance, the production of food-safe,
    non-denatured alcohol distilled from grain[^4] is accounted as
    output of the *Denatured Alcohol Manufacturing Industry* (NAICS[^5]
    325193, [@naicsco]). Grain remaining after distillation is a
    co-product[^6] product of denatured alcohol manufacturing and could
    be classified as a product of the *Other Animal Food Manufacturing
    Industry* (NAICS 311119, [@naicscoa]). One single process, two
    separate products.

3.  An economy is understood as an open system that can gain or lose
    income or products from outside its boundaries through trade.
    Moreover, any aggregations such as firms, cities, geopolitical
    regions, countries, or the entire world could form the boundaries of
    an economy examined. As described the *Encyclopedia of Social
    Measurement* [@vanleeuwen2005], large, diversified economies
    producing many goods and services can demonstrate large economic
    impacts because (a) a large region can fulfill its demand for goods
    and services from within the region, (b) transportation of goods and
    services with the region is simpler than hauling in these resources,
    and (c) economic activity in large region tends to agglomerate sales
    to surrounding regions. For these reasons, specification of the kind
    and level of aggregation of industry data can have a profound effect
    on an assessment of economic impact.

[^3]: If employees normally do not work at a single physical location,
    the establishment is identified by the main office that supervises
    employees' activities or is the base at which employees carry out
    these activities. 

[^4]: Wheat, barley, corn, sorghum, or cereals containing 55%-70%
    starch.

[^5]: *The North American Industry Classification System* (NAICS) is the
    standard used by Federal statistical agencies in classifying
    business establishments for the purpose of collecting, analyzing,
    and publishing statistical data related to the U.S. business economy
    [@northam].

[^6]: A *co-product*, a desirable secondary good that is generated
    during the manufacturing process and can be sold or reused
    profitably, is different than a *by-product*, a material of value
    that is produced as a residual of, or incidental to, the production
    process. Generally, co-products generate revenue; by-products are
    associated with costs.

*Linear*

Neither economies nor diseconomies of scale occur as producing industry
$j$ output increases or decreases (i.e., *linear*). Also, this
assumption implies that scale -- and the institutional knowledge
aggregated from collective work -- has no bearing on the dollar value of
resources required for production or for the functioning of the product
market.

##### Are These Assumptions Realistic?

How correct are these assumptions? How important even is verisimilitude
of the assumptions with reality?

These three assumptions -- fixed, homogeneous, linear -- about the
nature of direct input-output requirements are, of course, highly
restrictive and probably are unrealistic attributes to assign to most
economies. In spite of that, many analysts choose to believe that these
assumptions have proved tenable within limits and useful for economic
analysis in static contexts, even without adding additional complexity
to expand the simple model represented by equation (2). Rather, the use
of the model often rests on the verisimilitude of its assumptions with
reality. Many analysts have relied on the pragmatic view that these
assumptions are close enough to short-term reality to allow answers to
economic problems that do not lend themselves to other analytic
techniques [@goldman1965].

The John Locke Foundation is one critic of common input-output modeling
applications. The Foundation asserts [@economic2017] that many modeling
efforts fail to consider *opportunity costs* -- that is, forgone
benefits that could accrue from an option not chosen. Most studies using
input-output analysis only consider benefits of a project, but fail to
consider or model the costs of a project.

Of course, any new economic activity viewed without costs always will
demonstrate unalloyed benefits. Something is larger than nothing, right?
The "big-benefits-with-few-costs" approach is quite compelling to
legislators, policy-makers, and funding agencies seeking to convince the
public that projects, events, or changes are indispensable for economic
growth and community revitalization [@sanders2015]. The John Locke
Foundation asserts that input-output modeling is so often conducted "by
professional consulting firms who are hired by interest groups to do the
studies. Furthermore, seldom do those who actually perform the studies
have formal training in economics. Instead their expertise is in
using...proprietary [input-output] models [@economic2017a]".

Ignoring opportunity costs and other modes of failure to think ahead to
consider full economic consequences prompted Frédéric Bastiat, French
economist best known for his journalistic writing in favor of free trade
and the economics of Adam Smith, wrote in his last published pamphlet in
1850 that:

> In the economic sphere an act, a habit, an institution, a law produces
> not only one effect, but a series of effects. Of these effects, the
> first alone is immediate; it appears simultaneously with its cause; it
> is seen. The other effects emerge only subsequently; they are not
> seen; we are fortunate if we foresee them.

> There is only one difference between a bad economist and a good one:
> the bad economist confines himself to the visible effect; the good
> economist takes into account both the effect that can be seen and
> those effects that must be foreseen [@bastiat1964].

I do not mean to ignore these importance of these assumptions and
criticisms of them. I will take up these matters directly as I review
reports of economic impact in other postings on this blog.

#### Indirect and, then, Total Requirements (The Leontief Inverse)

Introducing for mathematical convenience an $I x J$ identity matrix,
$\textbf{I}$, and algebraically rearranging terms in equation (2), the
relationships among production, consumption and economic output is
expressed as

$$
\begin{equation}
\textbf{x}=(\textbf{I} - \textbf{A})^{-1}\textbf{y}, (3)
\end{equation}
$$

where $(\textbf{I} - \textbf{A})^{-1}$ is called a *total requirements
matrix* that shows the additional dollar value of output from industry
$i$ needed directly or indirectly to deliver a dollar's worth of output
from industry $j$ to final demand in the economy. The matrix inverse,
$(\textbf{I} - \textbf{A})^{-1}$, often is described as the *Leontief
Inverse*,[^7] in honor of Wassily Leontief [@leontief1936a;
@leontief1966c], an economist who won the Nobel Memorial Prize in
Economic Sciences in 1973 for developing economic input--output analysis
theory and methods [@thesver]. Actual economies can have large numbers
of industries, so $\textbf{A}$ often is highly dimensional. As a
consequence, $(\textbf{I} - \textbf{A})^{-1}$ usually is solved
numerically by computational techniques involving Laplace transforms and
Lagrange multipliers.

[^7]: ^-1^ refers to an inverse of a matrix, not a reciprocal.

Taking the difference between the total requirements matrix and the
direct requirements matrix produces an *indirect requirements* matrix.
Indirect requirements are generated from the long string of transactions
among industries representing intermediate production that must occur to
generate a final product for consumption. The addition of households as
both a row and column in $\textbf{X}$ allows inclusion of the effects of
household production and purchasing into the model. This addition
augments the total requirements matrix to include not only direct and
indirect requirements but also to show induced effects of household
spending on the economy as it stimulates purchases of goods and
services. The difference between the total requirements matrices, net of
the indirect requirements matrix, including household spending produces
an *induced requirements* matrix.

#### Multipliers

Indirect and induced requirements matrices display how each dollar spent
on goods and services delivered to consumers multiplies itself
throughout the economy. Ratios between [(direct requirements $+$
indirect requirements) $/$ direct requirements] as well as ratios
between [(direct requirements $+$ indirect requirements $+$ induced
requirements) $/$ direct requirements] are called *Type I* and *Type II*
multipliers, respectively. Pre--multiplying the direct, indirect, and
induced requirements matrices by, say, total employment parses the
impact of delivery of final demand to consumers into direct, indirect,
and induced employment components. In a similar manner, multipliers for
other economic indicators -- such as economic output, value added,
income, taxes -- can be computed.

People who commission input-output modeling of potential economic impact
often are most interested in multipliers to demonstrate that an
investment of public funds in a project can benefit an entire community.
For example, an analysis of the potential economic impact of a
petrochemical manufacturing facility in Beaver County (north of
Pittsburgh, PA) concludes

> ...operations will involve approximately 240 to 450 new jobs for
> Beaver County residents at the petrochemical facility. With
> multipliers, total new jobs held by Beaver County residents are
> anticipated to be between 777 and 1,444. Annual labor income increases
> in Beaver County will be between approximately \$73 and \$120 million.
> Over the 40-year operational life of the Project, labor income
> increases within Beaver County are estimated to total between \$1.5
> and \$2.4 billion. Annual value added in Beaver County would be
> between \$260 and \$846 million. Over the assumed 40-year operation
> period, this equates to a present value of \$10.3 to \$16.7
> billion.[@clinton2021a] 

This analysis and other assertions about economic impact of operation of
the petrochemical facility supported state and local tax breaks for the
developer:

> Pennsylvania's fiscal incentive package --- which includes a large tax
> credit along with job-training incentives --- was also crucial to the
> company's decision to both proceed with the project and locate it in
> Pennsylvania, Visser said.
>
> In 2012, state lawmakers authorized a tax credit for companies that
> use ethane that invest more than \$1 billion and employ more than
> 2,500 workers during construction. Shell was given a 25-year tax break
> worth about \$66 million a year starting in 2017, or about \$1.65
> billion in total. The state also extended a tax-free Keystone
> Opportunity Zone in Beaver County and provided other grants.
>
> "I can tell you with the hand on my heart that without these fiscal
> incentives, we would not have taken this investment decision," Visser
> [Royal Dutch Shell representative] said. [@location]

### Use of Counterfactuals to Identify Potential Economic Impact

#### What is a Counterfactual?

Input-output modeling is useful for describing of the anatomy and
physiology of an economy. However, these models also offer a basis for
simulating economic scenarios that are "counter to the facts" to inform
risk assessment.

A *counterfactual* is a hypothetical scenario that runs "counter to the
facts" of an existing reality. A counterfactual condition entails "What
if...?" questions that represent thought experiments about propositions
that run counter to the facts. For instance, how would an economy have
grown and changed if an activity or event did not occur? How do supply
chain constraints or disruptions affect an innovation? The potential
impact of an activity, event, or constraint is the difference between
what would have happened with and, then, without the activity, event, or
constraint.

Researchers long have used the counterfactual approach to consider
strategic decisions, decisive acts, and paths of activity that would
have altered the course of history. For instance, in *Ab Urbe Condita
Libri* [@aburbe1873a],[^8] Titus Livius Patavinus, a Roman historian,
contemplated an alternative in 4th century BC by questioning, "What
would have been the results for Rome if she had been engaged in war with
Alexander?" Lewis [@lewis1974] indicated that the modern counterfactual
approach goes back at least to the 18th century Scottish Enlightenment
philosopher David Hume.

[^8]: Book digitized by Google from the library of the University of
    Michigan and uploaded to the Internet Archive by user tpb.; 5
    volumes 20 cm; Pars 1. liber I-VI. -- pars 2. liber VII-XXIII. --
    pars 3. liber XXIV-XXX. -- pars 4. liber XXXI-XXXVIII. -- pars 5.
    liber XXXIX-XLV. Epitom. lib. XLVI-CXL. -- pars 6. Fragmenta et
    Index

Historians, essayists, and novelists have applied counterfactual
thinking productively in a variety of situations. For instance, in 1931
Winston Churchill wrote an essay, "If Lee Had Not Won the Battle of
Gettysburg," [@churchill2014] that examined the counterfactual viewpoint
about what would have happened if the Confederacy had won the American
Civil War[^9] (see counterfactual examples regarding the American Civil
War [@lewis1931]). In 1962, Philip K. Dick, a novelist in the science
fiction genre, published *The Man in the High Castle* [@dick1962], a
narrative of a thought experiment that considered the consequences if
Nazi Germany and Imperial Japan had won the Second World War.

[^9]: If Churchill's title seems like a non sequitur, refer to the
    explanation of his rhetoric at
    <http://www.troynovant.com/Franson/Churchill/If-Lee-Not-Won-Gettysburg.html>.

#### Constructing Counterfactuals Within Input-Output Modeling

Input--output modeling can help examine counterfactuals. The typical way
to conduct such a counterfactual approach is to construct and compare
two economic scenarios. The first scenario contains a possible
alteration in the economy -- the counterfactual -- that would be
identified in input-output terms an economic changes resulting from,
say, an innovation (e.g., a new production process), event (e.g., a
disaster), policy (e.g. immigration regulation), or financial rule
(e.g., taxing authority). A second scenario merely maintains the "status
quo" of the economy under the assumption that the economy simply grows
and changes consistently according to historical trends. The difference
between the first and the second scenarios is the potential economic
impact of the economic alteration specified in the first scenario.

The results of such counterfactual economic analyses are not economic
forecasts. Rather, such analyses isolate the importance only of one
narrow group of changes in an economy by showing the gross effect of
these changes rather that the effect of changes net of all other
potential changes taken together that could affect an economy. A
counterfactual used with input--output methods typically is specified
under the assumption of *ceteris paribus* -- that is, with the
assumption of "all other things remaining the same" or, as is much more
restrictively used in physics and other sciences [@cartwright1980], "if,
and only if, all other things are the same *and* correct" More
restrictively, *ceteris absentibus*, a term meaning literally "others
absent," probably best describes the approach for economic
counterfactuals [@joseph1980].

#### An Example of a Counterfactual Specification

A potential impact of a counterfactual is computed by calculating the
difference between (a) an economic scenario created by a counterfactual
condition exemplifying a production or consumption shock and (b) a
status quo scenario representing an economy anticipated under historical
trends in growth and change. Shown in Table 1 are three classes of
counterfactuals that are possible in a hypothetical case for the
production and consumption of *transparent aluminum*, a material I
introduce only to offer an example of the structure of potential
counterfactuals in input-output analysis.

Table 1. *Three hypothetical macroeconomic counterfactuals specified to
assess risk to production and consumption of transparent aluminum*

| **Counterfactual description**                                       | **Change specified in macroeconomic model**                                                                                                                                                                            |
|------------------------------------|------------------------------------|
| Change in consumption of transparent aluminum                        | $\Delta y_{i'}$, where element $y_{i'}$, is total final demand for industry $i'$, the transparent aluminum industry (where $i'\subseteq I$).                                                                           |
| Change in production function for the transparent aluminum industry  | $\Delta a_{j'}$, where vector $a_{j'}$ is a column vector representing the proportion of inputs from $I$ industries purchased as inputs to the transparent aluminum industry (where $j'\subseteq J$).                  |
| Change in production function of a purchaser of transparent aluminum | $\Delta a_{j}$, where vector $a_{j}$ is a column vector representing the proportion of inputs from $I$ industries purchased as inputs to any industry *other than* the transparent aluminum industry (where $j' ≠ j$). |

Transparent aluminum, a ceramic composite technically known as *aluminum
magnesium oxynitride*, is an extremely durable, sintered crystalline
material composed of magnesium, aluminum, oxygen, and nitrogen
(*AlN-MgO-MgAl2O4*) that is optically transparent (≥80%) in the
near-ultraviolet, visible, and midwave-infrared regions of the
electromagnetic spectrum. The name of this material comes from its
similarity to a material referenced in the science fiction movie, *Star
Trek IV: The Voyage Home*.[^10]

[^10]: Kirk and the Enterprise crew must save the whales to, in turn,
    save planet Earth. They must contain the whales in water on the
    Enterprise. The crew travels back in time to 20th Century Earth,
    when transparent aluminum had not been invented yet. Scott, the
    Enterprise's chief engineer, shares the formula for transparent
    aluminum with a 20th Century scientist so that the Enterprise could
    be outfit with the material crucial to building tanks for the
    whales.I dunno...it's supposed to be one of the light-hearted *Star
    Trek* movies.

Transparent aluminum is a promising material that is four times harder
than fused silica glass, 85% as hard as sapphire, and resists corrosion,
radiation, and oxidation [@nishiyama2017]. Products requiring both
durability and transparency -- such as armored glass, windshields on
airplanes and sea vessels, windows for laser and infrared
communications, personal protective gear, smartphone and watch screens,
and sensor components -- are among the potential uses for transparent
aluminum.

##### Change in consumption of transparent aluminum

Using equation (3), implementing this counterfactual involves
substitution of $y_{i'}$ for $y_{i}$ and solving for
$(\textbf{x}=(\textbf{I} - \textbf{A})^{-1}\textbf{y'}$, where $y'$ is a
new column vector of total final demand for industry $i'$, the
transparent aluminum industry ($i'\subseteq I$). The result is a direct
change in transparent aluminum output, along with indirect and induced
changes of the total output of all industries to the extent that they
have direct or indirect interindustry linkages to the transparent
aluminum industry.

##### Change in production function for the transparent aluminum industry

Changing a production function involves altering the proportion of total
products from other industries purchased by the transparent aluminium
industry. This change could result from efficiencies affected in the
production of transparent aluminium. Or, supply chain constraints or
shortages might have required changes in the manner that transparent
aluminium is produced. Such a change is manifest in alterations in
$\textbf{A}$. As a result, $(\textbf{x}=(\textbf{I} - \textbf{A})^{-1}$
also is altered, reestablishing the input-output identity to
$(\textbf{x}=(\textbf{I} - \textbf{A'})^{-1}\textbf{y}$ where
$\textbf{A'}$ is $\textbf{A}$ altered by $a_{j'}$, a column vector
representing the proportion of inputs from $I$ industries purchased as
inputs to the transparent aluminum industry ($j'\subseteq J$). The
result observed, holding transparent aluminum output constant, is
indirect and induced changes of the total output of all industries to
the extent that they have direct or indirect interindustry linkages to
the transparent aluminum industry.

##### Change in production function of a purchaser of transparent aluminium

The consequence of this scenario is a similar change similar to changes
in the production function for the transparent aluminum industry.
However, the result observed, holding output constant in the industry
making the production function change, is indirect and induced changes
of the total output of all industries, including the transparent
aluminum industry, to the extent that they have direct or indirect
interindustry linkages to the industry making the production function
change.

### Selected Software and Systems for Input-Output Modeling

Many software and systems are available to analyze input-output data
from the national income and product accounts of the United States (viz
description of accounts by the Bureau of Economic Analysis of the U.S.
Department of Commerce [@anintro2007]. Three prominent sets of software
and systems are *IMPLAN*, *RIMS II*, and *IO-Snap*.

#### IMPLAN

[IMPLAN](https://implan.com) (*IM*pact for *PL*anning and *AN*alysis) is
a prominent example of software used estimate the direct, indirect, and
induced impacts in regional economies. The IMPLAN economic impact
modeling system consists of software tools and region--specific data to
enable in--depth examinations of state, multi--county, county,
sub--county, and metropolitan regional economies. IMPLAN seems to have a
large share of the market for input-output modeling. Some writers (e.g.,
[@vowels2012a] ) describe IMPLAN as the "gold standard" for input-output
analysis.

#### RIMS II

[RIMS II](https://www.bea.gov/resources/methodologies/RIMSII-user-guide)
is a product of the Bureau of Economic Analysis of the U.S. Department
of Commerce that is based on a set of national input-output accounts
showing the goods and services produced by each industry and the use of
these goods and services by industries and final users.

#### IO-Snap

[IO-Snap](https://www.io-snap.com/) is a software application that
includes all aspects of fundamental input-output analyses. Data can be
edited and displayed in a variety of formats and compared and contrasted
across geographical definitions at different times.

## Last updated on {.appendix}

```{r,echo=FALSE}
Sys.time()
```

## Reuse {.appendix}

Text and figures are licensed under Creative Commons Attribution [CC BY
4.0](https://creativecommons.org/licenses/by-sa/4.0/). Source code is
available at <https://github.com/davidpassmore/blog>, unless otherwise
noted. The figures that have been reused from other sources don't fall
under this license and can be recognized by a note in their caption:
"Figure from ...".

## Comments/Corrections {.appendix}

To make comments about this posting or to suggest changes or
corrections, send email to [David Passmore](dlp@davidpassmore.net), send
a direct message on Twitter `@DLPPassmore`, or send an IMsg or SMS to
[dlp\@psu.edu](mailto:dlp@psu.edu){.email}.
