---
title: "can an analysis of economic impact have impact?"
description: |
  xxx
preview: xxx.jpg
author:
  - name: David L. Passmore
    url: https://davidpassmore.net
    orcid_id: 0000-0001-9307-5164
    affiliation: Distinguished Professor, Emeritus, Penn State; Academic Visitor, University of Pittsburgh
categories:
  - IMPLAN
  - Wassily Leontief
  - economic impact
  - counterfactuals
  - misspecification
  - identifying industry
draft: true
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

A regional macroeconomic model called IMPLAN (Impact for PLanning and
ANalysis) was applied to estimate the direct, indirect, and induced
impacts of egg production on the Pennsylvania economy. MIG, Inc., from
Hudson, Wisconsin, is the sole--source provider of the IMPLAN economic
impact modeling system, which consist of software tools and
region--specific data to enable in--depth examinations of state,
multi--county, county, sub--county, and metropolitan regional economies.
MIG, Inc. has been developing complex localized databases, conducting
IMPLAN training workshops, and distributing IMPLAN software to public
and private organizations since 1993. A copy of the IMPLAN software and
data package owned by Penn State's Institute for Research in Training
and Development was used in this study .

Methods for calculating economic multipliers are well--established and
common in the practice of economic impact analysis. However, in brief,

### *Appendix:* Input-Ouput Modeling in a Nutshell

An input-output model accounts for goods and services that are required
as *inputs* by industries to produce *outputs* of goods and services
necessary to satisfy supply requirements of other industries and for
delivery directly to consumers [@miernyk1965].

#### Specification of the Input-Output Model

##### Structure

Consider an economy with *I* producing industries and *J* purchasing
industries, where *I* = *J*. Some of the output of industry *i* is sold
to other *j* industries for use in their own production.

Let **X** stand for a square matrix of interindustry transactions, with
elements *x~ij~* containing the dollar value of goods and services sold
by producing industry *i* to purchasing industry *j*. Sales of the
product of industry *i* to itself --- that is, *x~i = j~* --- are
possible. For instance, the coal industry also might purchase coal for
use in its own production processes.

In this conception of an economy, industry output is produced by firms.
A number of firms can comprise an industry. Firms might own or manage
multiple establishments at which industry production occurs. To add
complexity, some firms produce output of goods and services that are
classified in multiple industries, which is, then, accounted separately
for each industry. The fundamental rule is that accounting for economic
output is by industry, which is an aggregation of highly similar good or
services, and not by firm or within establishments run by firms. Also,
the macroeconomy is understood as an open system in the sense that it
can gain or lose income or products from outside its boundaries through
trade. Moreover, any aggregations such as firms, cities, geopolitical
regions, countries, or the entire world could form the boundaries of the
macroeconomy examined.

Output of industry *i* not purchased by industries for their own
production instead is sold directly to consumers to meet demands for
personal consumption in households, government purchases of defense and
non-defense goods and services, fixed investments such as housing and
other structures, and exports net of imports. Let **y** indicate an
**I**--length vector whose elements display the dollar value of goods
and services delivered by producing industry *i* to fulfill final
demands of the economy for personal consumption, government purchases,
fixed investments, and net exports. Vector **y** represents the "final
demand" for goods and services in an economy and is known commonly in
the parlance of national income and product accounting practice as
*gross domestic product* [@grossdo].

##### Direct Requirements

Quantities in **X**, **y**, and **x** are most commonly expressed as
monetary values (e.g., as dollars, rubles, or some other form of
currency used as a medium of exchange) rather than as quantities of
goods and services, although some energy and resource input--output
models are scaled in physical quantities such as megawatts of
electricity, kilograms of waste, hectares of land, or liters of water
[@weisz2006].

The production sector of the economy, then is defined by **X**, and
**y** is the consumption sector. The total output of the economy by
producing industry, contained in *i*--length vector **x**,[^1] is equal
to the sum of production and consumption, or

[^1]: Although the notation seems awkward to have an element of matrix
    **X** indicated by the scalar x and the column vector of total
    output indicated by **x**, we are following notation conventions
    that are common in the field of interindustry analysis.

**x** = **X** + **y**. (1)

Define **A**, a direct requirements matrix with elements *a~ij~* =
*x~ij~* / *x~j~,* showing the proportion of industry *i* output required
for production by industry *j*. Equation 1 now becomes

**x** = **Ax** + **y**. (2)

Elements in **A** are assumed to be fixed, homogenous, and linear within
purchasing industries. In particular:

-   Inputs required by purchasing industry j from producing industry i
    do not change over time, or, more technically, zero elasticity of
    price or quantity substitution exists among inputs to purchasing
    industry j (i.e., *fixed*); \n
-   All firms in purchasing industry j use the same proportion of
    industry i inputs (i.e., *homogeneous*); \n
-   Neither economies nor diseconomies of scale are manifest as
    producing industry j output increases or decreases (i.e., *linear*).

These three assumptions are, of course, highly restrictive and probably
are unrealistic in a longer view of most economies. Yet, even without
adding additional complexity to detail the simple model represented by
equation (2), these assumptions have proved tenable within limits and
useful for economic analysis in static contexts. However, the use of the
model does not rest on the full verisimilitude of its assumptions.
Rather, many analysts have relied on the pragmatic view that these
assumptions are close enough to short-term reality to allow answers to
economic problems that do not lend themselves to other analytic
techniques [@goldman1965]

##### Indirect and, then, Total Requirements (The Leontief Inverse)

Introducing for mathematical convenience an *I*--by--*J* identity
matrix, **I**, and algebraically rearranging terms in equation 2, the
relationships among production, consumption and economic output is
expressed as

**x** = (**I** -- **A**)^-1^ **y**,

where (**I** -- **A**)^-1^ is called a *total requirements matrix*
showing the additional dollar value of output from industry *i* needed
directly or indirectly to deliver a dollar's worth of output from
industry *j* to final demand in the economy. The matrix inverse, (**I**
-- **A**)^-1^, often is described as the *Leontief Inverse*, in honor of
Wassily Leontief [@leontief1936a; @leontief1966c], an economist who won
the Nobel Memorial Prize in Economic Sciences in 1973 for developing
economic input--output analysis theory and methods [@thesver]. Actual
economies can have large numbers of industries, so **A** often has a
high number of dimensions. As a consequence, (**I** -- **A**)^-1^
usually is solved numerically by Laplace transform techniques.

Taking the difference between the total requirements matrix and the
direct requirements matrix produces an *indirect requirements* matrix.
Indirect requirements are generated from the long string of transactions
among industries representing intermediate production that must occur to
generate a final product for consumption. The addition of households as
both a row and column in X allows inclusion of the effects of household
production and purchasing into the model. This addition augments the
total requirements matrix to include not only direct and indirect
requirements but also to show induced effects of household spending on
the economy as it stimulates purchases of goods and services. The
difference between the total requirements matrices, net of the indirect
requirements matrix, including household spending produces an *induced
requirements* matrix.

##### Multipliers

Indirect and induced requirements matrices display how each dollar spent
on goods and services delivered to consumers multiplies itself
throughout the economy. Ratios between [(direct requirements + indirect
requirements) / direct requirements] as well as ratios between [(direct
requirements + indirect requirements + induced requirements) / direct
requirements] are called *Type I* and *Type II* multipliers,
respectively. Pre--multiplying the direct, indirect, and induced
requirements matrices by, say, total employment parses the impact of
delivery of final demand to consumers into direct, indirect, and induced
employment components. In a similar manner, multipliers for other
economic indicators -- such as economic output, value added, income --
can be computed.

#### Consideration of Counterfactuals with the Input-Output Model

##### What is a Counterfactual?

Input-output modeling is useful for describing of the anatomy and
physiology of an economy. However, these models also offer a basis for
simulating economic counterfactuals that can assist with risk
assessment. A *counterfactual* is a hypothetical scenario that runs
"counter to the facts" of an existing reality. A counterfactual
condition entails "What if...?" questions that represent thought
experiments about propositions that run counter to the facts. For
instance, how would an economy have grown and changed if an activity or
event did not occur? How do supply chain constraints or disruptions
affect an innovation? The potential impact of an activity, event, or
constraint is the difference between what would have happened with and,
then, without the activity, event, or constraint.

Researchers long have used the counterfactual approach to consider
strategic decisions, decisive acts, and paths of activity as if they had
not been executed in history. For instance, in *Ab Urbe Condita Libri*
[@aburbe1873a],[^2] Titus Livius Patavinus, a Roman historian,
contemplated an alternative in 4th century BC that questioned, "What
would have been the results for Rome if she had been engaged in war with
Alexander?" Lewis [@lewis1974]indicated that the modern counterfactual
approach goes back at least to the 18th century Scottish Enlightenment
philosopher David Hume. Historians, essayists, and novelists have
applied counterfactual thinking productively in a variety of situations.
For instance, in 1931 Winston Churchill wrote an essay, "If Lee Had Not
Won the Battle of Gettysburg," [@churchill2014] that examined the
counterfactual viewpoint about what would have happened if the
Confederacy had won the American Civil War^[If Churchill’s title seems like a non sequitur, refer to the explanation of his rhetoric at http://www.troynovant.com/Franson/Churchill/If-Lee-Not-Won-Gettysburg.html.] (see counterfactual examples regarding Civil
War [@lewis1931]). In 1962, Philip K. Dick, a novelist in the science
fiction genre, published *The Man in the High Castle* [@dick1962], a
narrative of a thought experiment that considered the consequences if
Nazi Germany and Imperial Japan had won the Second World War.

[^2]: Book digitized by Google from the library of the University of
    Michigan and uploaded to the Internet Archive by user tpb.; 5
    volumes 20 cm; Pars 1. liber I-VI. -- pars 2. liber VII-XXIII. --
    pars 3. liber XXIV-XXX. -- pars 4. liber XXXI-XXXVIII. -- pars 5.
    liber XXXIX-XLV. Epitom. lib. XLVI-CXL. -- pars 6. Fragmenta et
    Index

*****

##### Use of Counterfactuals Within Input-Output Modeling

Counterfactuals can be examined with input--output models. The typical
way to conduct such a counterfactual approach is to construct and
compare two economic scenarios. The first scenario contains a possible
alteration in the economy -- the counterfactual -- that would be
pertinent to realization of a structural innovation. For instance, an
alteration might result from changes in production of a strategic
material, such as steel, that is a requirement necessary to sustain a
structural innovation. The second scenario merely maintains the "status
quo" scenario under the assumption that the economy simply grows and
changes consistently according to historical trends. The difference
between the first and the second scenarios is the potential economic
impact of the economic alteration specified in the first scenario (e.g.,
the impact of decreased availability of steel on the innovation).

The results of such counterfactual economic analyses are not economic
forecasts. Rather, such analyses isolate the importance only of one
narrow group of changes in an economy by showing the gross effect of
these changes rather that the effect of changes net of all other
potential changes taken together that could affect an economy. A
counterfactual used with input--output methods typically is specified
under the assumption of *ceteris paribus* -- that is, with the
assumption of "all other things remaining the same" or, as is much more
restrictively used in physics and other sciences [8], "if, and only if,
all other things are the same *and* correct" In this vein, *Ceteris
absentibus*, a term meaning literally "others absent," probably best
describes the approach for economic counterfactuals [9].

A potential impact of a counterfactual is computed by calculating the
difference between (a) an economic scenario created by a counterfactual
condition exemplifying a production or consumption shock and (b) a
status quo scenario representing an economy anticipated under historical
trends in growth and change. Shown in Table 1 are three classes of
counterfactuals that are possible in a hypothetical case for the
production and consumption of *transparent aluminum*, a material I
introduce only to offer an example of the structure of potentail
counterfactuals in input-output analysis.

Transparent aluminum, a ceramic composite technically known as *aluminum
magnesium oxynitride*, is an extremely durable, sintered crystalline
material composed of magnesium, aluminum, oxygen, and nitrogen
(*AlN-MgO-MgAl2O4*) that is optically transparent (≥80%) in the
near-ultraviolet, visible, and midwave-infrared regions of the
electromagnetic spectrum. The name of this material comes from its
similarity to a material referenced in the science fiction movie, *Star
Trek IV: The Voyage Home*. This promising material is four times harder
than fused silica glass, 85% as hard as sapphire, and resists corrosion,
radiation, and oxidation [@nishiyama2017]. Products requiring both
durability and transparency -- such as armored glass, windshields on
airplanes and sea vessels, windows for laser and infrared
communications, personal protective gear, smartphone and watch screens,
and sensor components -- are among the potential uses for transparent
aluminum.

For example, researchers used the counterfactual approach to consider
strategic decisions, decisive acts, and paths as if they had not been
executed in history. For instance, in Ab Urbe Condita Libri, Titus
Livius Patavinus, a Roman historian, contemplated an alternative 4th
century BC that considered, "What would have been the results for Rome
if she had been engaged in war with Alexander?"

The modern counterfactual approach goes back at least to the 18th
century philosopher David Hume, and historians, essayists, and novelists
have applied it productively in a variety of situations. For instance,
in 1931 Winston Churchill wrote an essay, "If Lee Had Not Won the Battle
of Gettysburg," that took the counterfactual viewpoint of what would
have happened if the Confederacy had won the American Civil War. In
1962, Philip K. Dick, a novelist in the science fiction genre, published
The Man in the High Castle, a thought experiment in which Nazi Germany
and Imperial Japan won World War II.

------------------------------------------------------------------------

Input-output analysis is a technique for macroeconomic analysis based on
the interdependencies between economic sectors or industries. The
technique is commonly used to estimate the impacts of positive or
negative economic shocks and analyze the ripple effects throughout an
economy. Data for input-output analysis are derived from a country's
national income and product accounts, which are available for most
countries with modern economies. Wassily Leontief, a Russian-born
economist, won the 1973 Nobel Prize in Economics for his development of
input-output analysis. I have completed over 125 regional economic
science studies that have integrated national, regional, and local
economies using input-output analysis.

------------------------------------------------------------------------

2.1. Structure Consider an economy with I producing industries and J
purchasing industries, where I = J. Some of the output of industry i is
sold to other j industries for use in their own production. Let X stand
for a square matrix of interindustry transactions, with elements xij
containing the monetary value of goods and services sold by producing
industry i to purchasing industry j. Sales of the product of industry i
to itself -- that is, xii -- are possible. For instance, the coal
industry also might purchase coal for use in its own production
processes. In this conception of an economy, industry output is produced
by firms. A number of firms can comprise an industry. Firms might own or
manage multiple establishments at which industry production occurs. To
add complexity, some firms produce output of goods and services that are
classified in multiple industries, which is, then, accounted separately
for each industry. The fundamental rule is that accounting for economic
output is by industry, which is an aggregation of highly similar good or
services, and not by firm or within establishments run by firms. Also,
the macroeconomy is understood as an open system in the sense that it
can gain or lose income or products from outside its boundaries through
trade. Moreover, any aggregations such as firms, cities, geopolitical
regions, countries, or the entire world could form the boundaries of the
macroeconomy examined. Further, output of industry i not purchased by
industries for their own production instead is sold directly to
consumers to meet demands for personal consumption in households,
government purchases of defense and non-defense goods and services,
fixed investments such as housing and other structures, and exports net
of imports. Let y indicate an I--length vector whose elements display
the monetary value of goods and services delivered by producing industry
i to fulfill demands of the economy. Vector y represents the "final
demand" for goods and services in an economy and is known commonly in
the parlance of national income and product accounting practice as
"gross domestic product" [2]. 2.2. Direct requirements Quantities in X,
y, and x are most commonly expressed as monetary values (e.g., as
dollars, rubles, or some other form of currency used as a medium of
exchange) rather than as quantities of goods and services, although some
energy and resource input--output models are scaled in physical
quantities such as megawatts of electricity, kilograms of waste,
hectares of land, or litres of water [10]. The production sector of the
economy, then, is defined by X, and y is the consumption sector. The
total output of the economy by producing industry, contained in
I--length vector, x, is equal to the sum of production and consumption,
or x = X + y. (1) Define A, a direct requirements matrix with elements
aij = xij / xi, showing the proportion of industry i output required for
production by industry j. Equation 1 now becomes X = Ax + y. (2)
Elements in A are assumed to be fixed, homogenous, and linear within
purchasing industries. In particular: • Inputs required by purchasing
industry j from producing industry i do not change over time, or, more
technically, zero elasticity of price or quantity substitution exists
among inputs to purchasing industry j (i.e., fixed); • All firms in
purchasing industry j use the same proportion of industry i inputs
(i.e., homogeneous); • Neither economies nor diseconomies of scale are
manifest as producing industry j output increases or decreases (i.e.,
linear). These three assumptions are, of course, highly restrictive and
probably are unrealistic in a longer view of most economies. Yet, even
without adding additional complexity to detail the simple model
represented by equation (2), these assumptions have proved tenable
within limits and useful for economic analysis in static contexts.
However, the use of the model does not rest on the full verisimilitude
of its assumptions. Rather, many analysts have relied on the pragmatic
view that these assumptions are close enough to short-term reality to
allow answers to economic problems that do not lend themselves to other
analytic techniques [11]. 2.3. Total requirements: The Leontief Inverse
Introducing for mathematical convenience an I--by--J identity matrix, I,
and algebraically rearranging terms in equation 2, the relationships
among production, consumption and economic output is expressed as x = (I
-- A)-1 y, (3) where (I -- A)-1 is a total requirements matrix, which
shows the additional dollar value of output from industry i needed
directly or indirectly to deliver a dollar's worth of output from
industry j to final demand in the economy. The total requirements
matrix, (I -- A)-1, often is referenced as the Leontief Inverse,
acquiring its name after its originator, Wassily Leontief [12, 13], who
won the 1973 Nobel Prize in Economic Sciences largely for the
development and application of this technique [14]. Actual economies can
have large numbers of industries, so A often is large. As a consequence,
(I -- A)-1 usually is solved numerically by Laplace transform
techniques. Taking the difference between the total requirements matrix
and the direct requirements matrix produces an indirect requirements
matrix. Indirect requirements are generated from the long string of
transactions among industries representing intermediate production that
must occur to generate a final product for consumption. The addition of
households as both a row and column in X allows inclusion of the effects
of household production and purchasing into the model. This addition
augments the total requirements matrix to include not only direct and
indirect requirements but also to show induced effects of household
spending on the economy as it stimulates purchases of goods and
services. The difference between the total requirements matrices, net of
the indirect requirements matrix, including household spending produces
an induced requirements matrix. 2.4. Multipliers Indirect and induced
requirements matrices show how each dollar spent on goods and services
delivered to consumers multiplies itself throughout the economy. Ratios
such as [(direct requirements + indirect requirements) / direct
requirements] as well as [(direct requirements + indirect requirements +
induced requirements) / direct requirements] are called Type I and Type
II multipliers, respectively [15] and are common metrics used to account
for system-wide change evoked from changes in particular sectors of an
economy. Pre--multiplying the direct, indirect, and induced requirements
matrices by, say, total employment parses the impact of delivery of
final demand to consumers into direct, indirect, and induced employment
components. In a similar manner, multipliers for other economic indexes,
such as income or taxes, can be computed.

## Last updated on {.appendix}

```{r,echo=FALSE}
Sys.time()
```

## Reuse {.appendix}

Text and figures are licensed under Creative Commons Attribution [CC BY
4.0](https://creativecommons.org/licenses/by-sa/4.0/). Source code is
available at <https://github.com/davidpassmore/blog>, unless otherwise
noted. The figures that have been reused from other sources don't fall
under this license and can be recognized by a note in their caption:
"Figure from ...".

## Comments/Corrections {.appendix}

To make comments about this posting or to suggest changes or
corrections, send email to [David Passmore](dlp@davidpassmore.net), send
a direct message on Twitter `@DLPPassmore`, or send an IMsg or SMS to
[dlp\@psu.edu](mailto:dlp@psu.edu){.email}.
