---
title: "xxx"
description: |
  xxx
preview: xxx.jpg
author:
  - name: David L. Passmore
    url: https://davidpassmore.net
    orcid_id: 0000-0001-9307-5164
    affiliation: Distinguished Professor, Emeritus, Penn State; Academic Visitor, University of Pittsburgh
date: 2022-03-28
categories:
  - xxxx
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web. 

Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.


## Last updated on {.appendix}
```{r,echo=FALSE}
Sys.time()
```
## Reuse {.appendix}

Text and figures are licensed under Creative Commons Attribution [CC BY 4.0](https://creativecommons.org/licenses/by-sa/4.0/). Source code is available at https://github.com/davidpassmore/blog, unless otherwise noted. The figures that have been reused from other sources don't fall under this license and can be recognized by a note in their caption: "Figure from ...".

## Comments/Corrections {.appendix}

To make comments about this posting or to suggest changes or corrections, send email to [David Passmore](dlp@davidpassmore.net), send a direct message on Twitter @DLPPassmore, or send an IMsg or SMS to dlp@psu.edu.




